Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-03-19 paper.qmd
Fixed Figure 2

@@

two_curvy_proj_true2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = true_model_df_proj, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#2b8cbe") +
  geom_point(
    size = 0.5,
    alpha = 0.2,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.6, 0.6)) + 
  ylim(c(-0.6, 0.6)) +
  interior_annotation("b", sc_ltr_pos, cex = 1.5) +
  theme(
    legend.position = "none"
  )

@@

```{r}
#| label: two-curvy-true-proj
# Apply the scaling
df_model_data_two_curvy <- bind_rows(data_two_curvy, true_model_two_curvy)

scaled_two_curvy_data_model <- scaled_two_curvy |>
  filter(type == "true model") |>
  select(-type)

proj_obj1 <- get_projection(projection = projection, 
               proj_scale = 1, 
               scaled_data = scaled_two_curvy_data, 
               scaled_data_model = scaled_two_curvy_data_model, 
               distance_df_small_edges = distance_df_small_edges_two_curvy, 
               axis_param = list(limits = 0.6, 
                                 axis_scaled = 2,
                                 axis_pos_x = -0.4, 
                                 axis_pos_y = -0.4, 
                                 threshold = 0.04))

# projection_scaled <- projection * 1
# 
# projected <- as.matrix(scaled_two_curvy_data) %*% projection_scaled
# 
# projected_df <- projected |>
#   tibble::as_tibble(.name_repair = "unique") |>
#   dplyr::rename(c("proj1" = "...1",
#                   "proj2" = "...2")) |>
#   #dplyr::mutate(type = df_exe$type) |>
#   dplyr::mutate(ID = dplyr::row_number()) 
# 
# projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled
# 
# projected_model_df <- projected_model |>
#   tibble::as_tibble(.name_repair = "unique") |>
#   dplyr::rename(c("proj1" = "...1",
#                   "proj2" = "...2")) |>
#   dplyr::mutate(ID = dplyr::row_number())
# 
# model_df <- dplyr::left_join(
#   wireframe_true_model, 
#   projected_model_df, 
#   by = c("from" = "ID"))
# 
# names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
# 
# model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
# names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
# 
# axes_obj <- gen_axes(  
#   proj = projection * 2,
#   limits = 0.6,
#   axis_pos_x = -0.4,
#   axis_pos_y = -0.4,
#   axis_labels = names(scaled_two_curvy_data),
#   threshold = 0.04)
# 
# axes <- axes_obj$axes
# circle <- axes_obj$circle

two_curvy_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#2b8cbe") +
  geom_point(
    size = 0.5,
    alpha = 0.5,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.6, 0.6)) + 
  ylim(c(-0.6, 0.6)) +
  interior_annotation("a", c(0.08, 0.93), cex = 1.5) +
  theme(
    legend.position = "none"
  )

proj_plot_two_curvy_true <- model_df |>
  ggplot(
    aes(
      x = proj1_from, 
      y = proj2_from)) +
  geom_point(
    color = "#000000",
    alpha = 0.5) +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.6, 0.6)) + 
  ylim(c(-0.6, 0.6)) +
  interior_annotation("a", c(0.08, 0.93), cex = 1.5) +
  theme(
    legend.position = "none"
  )

two_curvy_proj_umap_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#2b8cbe") +
  geom_point(
    size = 0.5,
    alpha = 0.5,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.42, 0.42)) + 
  ylim(c(-0.42, 0.42)) +
  interior_annotation("b", cex = 1.5) +
  theme(
    legend.position = "none"
  )
```

@@

two_curvy_proj_true + two_curvy_proj_true2 +
#two_curvy_proj_umap_model1 + two_curvy_proj_umap_model2 + 
  plot_layout(ncol = 2)
```

>>>>

projected_true_model <- as.matrix(scaled_two_curvy_data_true_model) %*% projection

@@

two_curvy_proj_true2 <- projected_df |>
  ggplot(
    aes(
      x = proj1,
      y = proj2)) +
  geom_segment(
    data = true_model_df_proj,
    aes(
      x = proj1_from,
      y = proj2_from,
      xend = proj1_to,
      yend = proj2_to),
    color = "#2b8cbe") +
  geom_point(
    size = 0.5,
    alpha = 0.2,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2),
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle,
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.42, 0.42)) +
  ylim(c(-0.42, 0.42)) +
  interior_annotation("b", sc_ltr_pos, cex = 1.5) +
  theme(
    legend.position = "none"
  )

@@

@@

two_curvy_proj_true + two_curvy_proj_true2 +
  plot_layout(ncol = 2)
```
